Study of the hadronization of charged pions

### 2019-03-13 intro.tex
Small fixes (pt max = 1.5 GeV2)
Minor text edits in introductory chapter

The hadronization in nuclear matter is a key process of the Quantum 
Chromo-Dynamics theory (QCD) because it leads to major theoretical 
uncertainties in numerous measurements. DIS experiments at JLab energies are 
well suited to test different theoretical models of hadronization, including 
the ones used for Relativistic Heavy Ion Collisions (RHIC). The main advantage 
of testing models in lepton scattering is that the size of cold nuclei is 
stable and known, while in RHIC the nuclear medium is hot and evolving 
introducing many parameters and unknowns. The hadronization process can also 
give important information about nuclei through the quark energy loss, for example, Arleo's model~\cite{Arleo:2003yf} correlates the quark energy loss with a gluon density, and Kopeliovich et al.'s model~\cite{Kopeliovich:2010aa} correlates the $\Delta \langle P_T^2 \rangle$ with the saturation scale. Other models \cite{Gallmeister:2007an} utilize different assumptions to describe the pre-hadron evolution into a full hadron. Therefore, high precision measurements are very important to discriminate between different models.

A quantitative understanding of the hadronization process in nuclei is also of importance for neutrino experiments. Indeed they use nuclear targets to maximize their cross sections and therefore have to account for the nuclear effects, including hadronization.

@@

In this analysis we use the following Semi-Inclusive Deep Inelastic Scattering (SIDIS) variables:
\begin{itemize}
 \item $\nu = E_i - E_f$ is the energy transferred by the lepton probe in the laboratory frame, where $E_i$ is the beam energy and $E_f$ is the scattered electron energy,
 \item $Q^2 = 4 E_i E_f \sin ^2(\theta_e / 2)$ is the 4-momentum transferred, with $\theta_e$ is the polar angle of a scattered electron,
 \item $x_{Bj} = {{Q^2} \over {2 M_n \nu}}$ is the proton momentum fraction carried by the struck quark, with $M_n$ is the nucleon mass,
 \item $W^2 = M_n^2 - Q^2 + 2 M_n \nu$ is the mass squared of the hadronic final state,
 \item $z_h = E_h / \nu$ is the virtual photon energy fraction carried by the measured hadron, with $E_h$ is the hadron energy,
 \item $P_T^2$ is the hadron transverse momentum measured with regard to the virtual photon direction,
 \item $\phi_h$ is the angle between the leptonic plane (containing initial and scattered electrons) and the hadronic plane (containing the virtual photon and a measured hadron),
 \item $x_F = P_L/P_L^{max}$ is the longitudinal momentum fraction carried by the hadron, and calculated with respect to the virtual photon direction in the laboratory frame.
\end{itemize}
%TODO Add t

We define the multiplicity ratio as:
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_T^2) = {{N_A^h (Q^2,\nu,z_h,P_T^2) / N_A^e (Q^2,\nu)} 
                       \over {N_D^h (Q^2,\nu,z_h,P_T^2) / N_D^e (Q^2,\nu)}},
\end{equation}
$N^e_A$ and $N_A^h$ are the number of electrons and semi-inclusive hadrons $h$ measured simultaneously on a target $A$. The multiplicity ratio represents the attenuation of the hadron $h$ in a nuclear target~$A$.
\newline

@@

The different models explaining the data can be separated in three families;
some assume that the quark looses energy in the medium and that either the hadronization occurs outside the medium or the hadronic interaction is 
negligible, others neglect the quark energy loss and consider
only the hadron and pre-hadron absorption, and the third category considers both
interactions. For illustration, we will give an example of each case, but for more detailed information, we refer the readers again to our reference~\cite{Accardi:2009qv}, which is more exhaustive and highlights also models describing RHIC's experiments.

In \cite{Wang:2002ri}, E.~Wang and X.-N.~Wang, described HERMES data using only 
the parton energy loss, hence the observed suppression is due to the fact that a lower energy quark fragments into a low number of hadrons at a very low $z$. This kind of model is suitable for both RHIC and nDIS experiments since it permits a common interpretation of hadron suppression in the nuclear matter. Moreover, various calculations have described the parton energy loss in the nuclear medium using the $\hat q$ (GeV$^2$ fm$^{-1}$) parameter, that is defined as the quark transverse momentum normalized by its propagation path length. In these models, $\hat q$ is mainly related to the $P_T$ broadening observable. The main difficulty of quark energy loss models is the lack of a coherent description for both multiplicity ratios and $P_T$ broadening behaviors. Recent models used a $\hat q$ to reproduce multiplicity ratio, which is too large to describe the transverse momentum broadening.
% TODO Add a ref

The other transport model, GiBUU~\cite{Gallmeister:2007an}, is based on the 
Boltzmann equation that employs hadronic and pre-hadronic interactions in the nuclear matter without involving any quark energy loss. This model reproduces very well most of hadrons' multiplicity ratios, however, it failed to describe the transverse momentum broadening. 
%$\Delta \langle P_T^2 \rangle$ variable is not described at all by

Furthermore, B.Z.~Kopeliovich et al. \cite{Kopeliovich:2008uy} described the process by neither neglecting the quark energy loss nor the hadron absorption, but used both $\Delta \langle P_T^2 \rangle$ and $R_A^h$ observables to differentiate between these effects. In that case, the transverse momentum broadening is correlated with the quark energy loss, and the suppression of multiplicity ratios is explained by the hadron absorption in the nuclear medium.

@@

Hadron multiplicity ratios in nuclei were measured in numerous lepton 
facilities: L.S.~Osborne {\it et al.} \cite{Osborne:1978ai} at SLAC,
L.~Hand {\it et al.} \cite{Hand:1978tx}, the E665 collaboration \cite{Adams:1994ri} at FNAL, and the European Muon Collaboration \cite{Arvidson:1984fz,Ashman:1991cx} at CERN. These measurements revealed the main features of the hadronization mechanism
in nuclei by showing a suppression of hadrons' production in heavy nuclei.
This suppression appears to be reduced at higher $\nu$ and lower $z$, at very 
low $z$ it can even change sign and become an increased yield.

Figures \ref{fig:her1}, \ref{fig:her2} and \ref{fig:her3} show a sample of 
the most recent HERMES data~\cite{Airapetian:2007vu}, where numerous hadrons were individually studied, and new variables linked with the transverse momentum were used in addition to the usual multiplicity ratio.

>>>>

The hadronization in nuclear matter is also a key process for the study of the Quantum 
Chromo-Dynamics theory (QCD) because it leads to major theoretical 
uncertainties in numerous measurements. DIS experiments at JLab energies are 
well suited to test different theoretical models of hadronization, including 
the ones used for Relativistic Heavy Ion Collisions (RHIC). The main advantage 
of testing models in lepton scattering is that the size of cold nuclei is 
stable and known, while in RHIC the nuclear medium is hot and evolving 
introducing many parameters and unknowns. The hadronization process can also 
give important information about nuclei through the quark energy loss, for example, Arleo's model~\cite{Arleo:2003yf} correlates the quark energy loss with a gluon density, and Kopeliovich et al.'s model~\cite{Kopeliovich:2010aa} correlates the $\Delta \langle P_T^2 \rangle$ with the saturation scale. Other models \cite{Gallmeister:2007an} utilize different assumptions to describe the pre-hadron evolution into a full hadron. Therefore, high precision measurements are very important to discriminate between different models.

A quantitative understanding of the hadronization process in nuclei is also of importance for neutrino experiments. Indeed, these use nuclear targets to maximize their cross sections and therefore have to account for the nuclear effects, including hadronization.

@@

In this analysis, we use the following Semi-Inclusive Deep Inelastic Scattering (SIDIS) variables:
\begin{itemize}
 \item $\nu = E_i - E_f$ is the energy transferred by the lepton probe in the laboratory frame, where $E_i$ is the beam energy and $E_f$ is the scattered electron energy,
 \item $Q^2 = 4 E_i E_f \sin ^2(\theta_e / 2)$ is the 4-momentum transferred, where $\theta_e$ is the polar angle of a scattered electron,
 \item $x_{Bj} = {{Q^2} \over {2 M_n \nu}}$ is the proton momentum fraction carried by the struck quark, with $M_n$ the nucleon mass,
 \item $W^2 = M_n^2 - Q^2 + 2 M_n \nu$ is the mass squared of the hadronic final state,
 \item $z_h = E_h / \nu$ is the virtual photon energy fraction carried by the measured hadron, with $E_h$ the hadron energy,
 \item $P_T^2$ is the hadron transverse momentum measured with regard to the virtual photon direction,
 \item $\phi_h$ is the angle between the leptonic plane (containing initial and scattered electrons) and the hadronic plane (containing the virtual photon and a measured hadron),
 \item $x_F = P_L/P_L^{max}$ is the longitudinal momentum fraction carried by the hadron, and calculated with respect to the virtual photon direction in the laboratory frame.
 Effectively this translate into: $$ x_F = 2 {z M_n \nu^2 - z Q^2 \nu - (M_n+\nu)(\boldsymbol{P_h \cdotp q}) \over (W^2-M_h^2) \sqrt{\nu^2+Q^2} } .$$
\end{itemize}
%TODO Add t

We define the multiplicity ratio as:
\begin{equation}
R_A^h (Q^2,\nu,z_h,P_T^2) = {{N_A^h (Q^2,\nu,z_h,P_T^2) / N_A^e (Q^2,\nu)} 
                       \over {N_D^h (Q^2,\nu,z_h,P_T^2) / N_D^e (Q^2,\nu)}},
\end{equation}
$N^e_A$ and $N_A^h$ are the number of electrons and semi-inclusive hadrons $h$ measured simultaneously on a target $A$. The multiplicity ratio represents the relative production rate of the hadron $h$ in a nuclear target~$A$.
\newline

@@

The different models explaining the data can be separated in three families;
some assume that the quark looses energy in the medium and that either the hadronization occurs outside the medium or the hadronic interaction is 
negligible, others neglect the quark energy loss and consider
only the hadron and pre-hadron absorption, and the third category considers both
interactions. For illustration, we will give an example of each case, but for more detailed information, we refer the readers again to the Accardi {\it et al.} review~\cite{Accardi:2009qv}, which is more exhaustive and highlights also models describing RHIC's experiments.

In \cite{Wang:2002ri}, E.~Wang and X.-N.~Wang, described HERMES data using only 
the parton energy loss, hence the observed suppression is due to the fact that a lower energy quark fragments into a low number of hadrons at a very low $z$. This kind of model is suitable for both RHIC and nDIS experiments since it permits a common interpretation of hadron suppression in the nuclear matter. Moreover, various calculations have described the parton energy loss in the nuclear medium using the $\hat q$ (GeV$^2$ fm$^{-1}$) parameter, that is defined as the quark transverse momentum normalized by its propagation path length. In these models, $\hat q$ can also be related to the $\Delta \langle P_T^2 \rangle$ observable. The main difficulty of quark energy loss models is the lack of a coherent description for both multiplicity ratios and $P_T$ broadening behaviors. Recent models used a $\hat q$ to reproduce multiplicity ratio, which is too large to describe the transverse momentum broadening.
% TODO Add a ref

The transport model, GiBUU~\cite{Gallmeister:2007an}, is based on the 
Boltzmann equation that employs hadronic and pre-hadronic interactions in the nuclear matter without involving any quark energy loss. This model reproduces very well most of hadrons' multiplicity ratios, however, it failed to describe the transverse momentum broadening. 
%$\Delta \langle P_T^2 \rangle$ variable is not described at all by

Finally, B.Z.~Kopeliovich et al. \cite{Kopeliovich:2008uy} described the process by neither neglecting the quark energy loss nor the hadron absorption, but used both $\Delta \langle P_T^2 \rangle$ and $R_A^h$ observables to differentiate between these effects. In that case, the transverse momentum broadening is correlated with the quark energy loss, and the suppression of multiplicity ratios is explained by the hadron absorption in the nuclear medium.

@@

Hadron multiplicity ratios in nuclei were originaly measured in numerous lepton 
facilities: L.S.~Osborne {\it et al.} \cite{Osborne:1978ai} at SLAC,
L.~Hand {\it et al.} \cite{Hand:1978tx}, the E665 collaboration \cite{Adams:1994ri} at FNAL, and the European Muon Collaboration \cite{Arvidson:1984fz,Ashman:1991cx} at CERN. These measurements revealed the main features of the hadronization mechanism
in nuclei by showing a suppression of hadrons' production in heavy nuclei.
This suppression appears to be reduced at higher $\nu$ and lower $z$, at very 
low $z$ it can even change sign and become an increased yield.

Figures \ref{fig:her1}, \ref{fig:her2} and \ref{fig:her3} show a sample of 
the most recent HERMES data~\cite{Airapetian:2007vu} from DESY, where numerous hadrons were individually studied, and new variables linked with the transverse momentum were used in addition to the traditional multiplicity ratio.
